---
title: "penguins_lab_6"
author: "Sam Csik"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. load packages and read in the penguins.csv data
packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archipelago?

Single factor = species
Levels = 3
Continuous random variable = flipper length

Maybe I can use a one-way ANOVA...

Assumptions of one-way ANOVA:
-Independent observations
-Equal variances (Levene's test)
-Normality (histograms/qqplots)

```{r}
counts <- with(penguins, table(Species, Sex))

penguins2 <- penguins %>% 
  mutate(
    common_name = # new column "common_name"
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo", 
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
    ) %>% 
  select(-Species) %>% 
  filter(Sex == "FEMALE")

```

Exploratory graphs and Levene's test for equal variances

```{r}
hists <- ggplot(penguins2, aes(x = `Flipper Length (mm)`)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")

qqs <- ggplot(penguins2, aes(sample = `Flipper Length (mm)`)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")

# based on histograms & qqplots, observations of each group appear normally distributed. Also, n > 30 for all groups, so means will be normally distributed by CLT.
```

Evaluate variances and perform a Levene's test for equal variances
```{r}
# if the largest sample is < 4x greater than the smallest sample variancer, thenathese are "close enough" to consider variances equal

variances <- penguins2 %>% # units of var are obs^2
  group_by(common_name) %>% 
  summarise(
    variance = var(`Flipper Length (mm)`)
  )

#largest var (33.1) is defiitely not more than 4x greater than the smallest variance (15.2)

# Levene's test
# leveneTest function in the 'car' package

penguin_levene <- leveneTest(`Flipper Length (mm)` ~ common_name, data = penguins2)

# H0: no difference in variances (variances are equal)
# Ha: variances are not equal
# we retain H0! there is no significant difference in variances across species
```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}
# H0: means across all groups are equal
# Ha: at least 2 group means differ significantly

penguin_aov <- aov(`Flipper Length (mm)` ~ common_name, data = penguins2)
summary(penguin_aov)

# post-hoc testing by Tukey's HSD
# H0: No sig diff in pairwise means
# Ha: there is a sig diff in pairwise means
penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph # reject H0 for ALL pairwise comparisons

# aov.test funciton in the 'onewaytests' package DOES allow for inline references
```
Mean flipper lengths of female penguins differed significantly for all species studied (one-way ANOVA, F(df = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).
