---
title: "penguins_lab_6"
author: "Sam Csik"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. load packages and read in the penguins.csv data
packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie penguins near Palmer Archipelago?

Single factor = species
Levels = 3
Continuous random variable = flipper length

Maybe I can use a one-way ANOVA...

Assumptions of one-way ANOVA:
-Independent observations
-Equal variances (Levene's test)
-Normality (histograms/qqplots)

```{r}
counts <- with(penguins, table(Species, Sex))

penguins2 <- penguins %>% 
  mutate(
    common_name = # new column "common_name"
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", 
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo", 
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>% 
  select(-Species) %>% 
  filter(Sex == "FEMALE")

```

